---
Title: "Endometrium driver mutation burden model"
author: "Luiza Moore"
date: "19 December 2018"
output: html_document
---

# Endometrium driver mutation burden modeling

Markdown file to document the methods used in analysing the mutation burden of endometerium.

## Load Libraries
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
library(rlang) 
library(knitr)
library(kableExtra)
library(pbkrtest)
```


## Load in data
Here, we load sample level data for 15/18 donors with data on BMI and parity.
```{r Load_data}
endom_burden <- read.csv("endom_burden_summary_file_for_model.final.csv")
# samples per patient
endom_burden %>% group_by(PatientID) %>%  count(PatientID) %>%  rename(`Sample count` = n) %>% arrange(desc(`Sample count`)) %>%  kable() %>%  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F, position = "left")
```


```{r Plot_1, fig.align='center'}
# Look at raw data
  endom_burden %>% ggplot(aes(Age, Total_drivers, colour = PatientID)) +
    geom_jitter(width = 0.2) +
    theme(plot.title = element_text(size = 8)) +
    ggtitle("Age-associated accumulation of driver mutations\nin human endometrium (substitutions only)") +
    theme(plot.title = element_text(size = 14)) + theme_bw() +theme(plot.title = element_text(hjust = 0.5))
```



## Fit model to estimate driver mutation rate per year
To account for the non-independent sampling per patient we use a generalized linear mixed effects model with Poisson distribution.
We also use a random slope with fixed intercept as most women will start menarche at a similar age (~13 years), but to account for the potential differences in the rates at which mutations were acquired in different individuals due to variation in parity, contraception and other factors. 

We test features with a known affect on mutation burden or endometrial cancer risk;

*Age

*Read depth & VAF 

*BMI

*Parity 

### Define full model and drop each fixed effect in turn
```{r Fit linear mixed effects models to estimate mutation rate per year, message= FALSE, warning=FALSE}
# Combine read depth and median sample depth as Vafdepth
endom_burden %<>%  mutate(Vafdepth = Seq_X*Sample_Median_Vaf)

full_glmer_model = glmer(Total_drivers ~ Age + Vafdepth + BMI + Parity + (Age - 1|PatientID), data=endom_burden, family = poisson(link = "log"), control =  glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

# "user" parametric boot function as defined in drop1.merMod help example
 PBSumFun <- function(object, objectDrop, ...) {
   pbnames <- c("stat", "p.value")
   r <- if (missing(objectDrop)) {
   setNames(rep(NA, length(pbnames)), pbnames)
   } else {
   # pbtest <- PBmodcomp(object, objectDrop, nsim = nsim, ref = NULL, seed=NULL, cl = clFit, details = 0)
    pbtest <- PBmodcomp(object, objectDrop, nsim = nsim, ref = NULL, seed=NULL, details = 0)
   unlist(pbtest$test[2, pbnames])
   }
   attr(r, "method") <- c("Parametric bootstrap via pbkrtest package")
   r
 }
 
nsim <- 1000
drop1(full_glmer_model, test = "user", sumFun = PBSumFun)
# stopCluster(clFit)
```

### Remove any feature with P > 0.05 to make final model
```{r Final model, message= FALSE, warning=FALSE}
# Remove Vafdepth and BMI from full model
final_glmer_model <- update(full_glmer_model, ~ . -BMI -Vafdepth, control=glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))
# Drop each fixed effect from final model
nsim = 1000
drop1(final_glmer_model, test = "user", sumFun = PBSumFun)
```
### Print Final model
```{r Print Final model}
# Print final model summary
print(summary(final_glmer_model))
```


