---
Title: "Endometrium total mutation burden model"
author: "LM"
date: "13 December 2018"
output: html_document
---

# Endometrium total mutation burden modeling

Markdown file to document the methods used in analysing the mutation burden of endometerium.

## Load Libraries
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
library(rlang) 
library(knitr)
library(kableExtra)
```


## Load in data
Here, we load sample level data for 15/18 donors with data on BMI and parity.
```{r Load_data}
endom_burden <- read.csv("endom_burden_summary_file_for_model.final.csv")
# samples per patient
endom_burden %>% group_by(PatientID) %>%  count(PatientID) %>%  rename(`Sample count` = n) %>% arrange(desc(`Sample count`)) %>%  kable() %>%  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F, position = "left")
```


```{r Plot_1, fig.align='center'}
# Look at raw data
endom_burden %>% ggplot(aes(Age, Subs, colour = PatientID)) +
  geom_jitter(width = 0.2) +
  theme(plot.title = element_text(size = 8)) +
  ggtitle("Age-associated accumulation of somatic mutations\nin human endometrium (substitutions only)") +
  theme(plot.title = element_text(size = 14)) + theme_bw() +theme(plot.title = element_text(hjust = 0.5))
```



## Fit linear mixed effects models and estimate mutation rate per year
To account for the non-independent sampling per patient we use a linear mixed effects model as the observed frequencies of all substitutions approximates a normal distribution.
We also use a random slope with fixed intercept as most women will start menarche at a similar age (~13 years), but to account for the potential differences in the rates at which mutations were acquired in different individuals due to variation in parity, contraception and other factors. 

We test features with a known affect on mutation burden or endometrial cancer risk;

* Age
* Read depth & VAF 
* Driver mutations
* BMI
* Parity 

We use backwards elimination to define a final model 

### Make full model and drop each fixed effect in turn
```{r Fit linear mixed effects models to estimate mutation rate per year}
# Combine read depth and median sample depth as Vafdepth
endom_burden %<>%  mutate(Vafdepth = Seq_X*Sample_Median_Vaf)

full_lmer_model = lmer(Subs ~ Age + Vafdepth + BMI + Parity + Driver_status + (Age - 1|PatientID), data=endom_burden, REML=F)
# Drop each fixed effect
lme4:::drop1.merMod(full_lmer_model, test = "Chisq")
```

### Remove feature with largest P > 0.05 to make reduced model 1
```{r Reduced model 1, message= FALSE, warning=FALSE}
# Remove parity from full model
reduced1_glmer_model <- update(full_lmer_model, ~ . -Parity)
# Drop each fixed effect from final model
lme4:::drop1.merMod(reduced1_glmer_model, test = "Chisq")
```
### Define Final model
```{r Print Final model}
# Define final model as all features are significant P < 0.05
final_glmer_model <- reduced1_glmer_model
# Print final model summary
print(summary(final_glmer_model))

