---
Title: "Endometrium mutation burden model"
author: "LM"
date: "13 December 2018"
output: html_document
---

# Endometrium mutation burden modeling

Markdown file to document the methods used in analysing the mutation burden of endometerium.

## Load Libraries
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
# library(readr)
# library(dplyr)
# library(ggplot2)
# library(readxl)
library(magrittr)
library(lme4)
library(lmerTest)
library(rlang) 
library(knitr)
library(kableExtra)
```


## Load in data
Here, we load sample level data for 15/18 donors with data on BMI and parity.
```{r Load_data}
 endom_burden <- read.csv("endom_burden_summary_file_for_model.final.csv")
  # head(endom_burden)
  endom_grouped <- group_by(endom_burden, PatientID)
  # head(endom_grouped)
  
  endom_grouped %>% group_by(PatientID) %>%  count(PatientID) %>%  rename(`Sample count` = n) %>% arrange(desc(`Sample count`)) %>%  kable() %>%  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F, position = "left")
```


```{r Plot_1, fig.align='center'}
    
# Look at raw data
  endom_burden %>% ggplot(aes(Age, Total_drivers, colour = PatientID)) +
    geom_jitter(width = 0.2) +
    theme(plot.title = element_text(size = 8)) +
    ggtitle("Age-associated accumulation of driver mutations\nin human endometrium (substitutions only)") +
    theme(plot.title = element_text(size = 14)) + theme_bw() +theme(plot.title = element_text(hjust = 0.5))
```



## Fit linear mixed effects models and estimate driver mutation rate per year
To account for the non-independant sampling per patient we use a linear mixed effects model.
We also use a random slope with fixed intercept as most women will start menarche at a similar age (~13 years), but to account for the potential differences in the rates at which mutations were acquired in different individuals due to variation in parity, contraception and other factors. 

We test features with a known affect on mutation burden or endometrial cancer risk;

*Age

*Sensitivty 

*BMI

*Parity 

### Make full model and drop each fixed effect in turn
```{r Fit linear mixed effects models to estimate mutation rate per year}
full_glmer_model = glmer(Total_drivers ~ Age + BMI + Parity + (Age - 1|PatientID), data=endom_burden, family = "poisson")
kable(lme4:::drop1.merMod(full_glmer_model)) %>%  kable_styling(bootstrap_options = c("striped", "condensed"))
```

### Remove any feature with an model AIC less than full model
```{r Final model}
# Remove BMI from full model
reduced_glmer_model <- update(full_glmer_model, ~ . -BMI)
kable(lme4:::drop1.merMod(reduced_glmer_model)) %>%  kable_styling(bootstrap_options = c("striped", "condensed"))

```
### Print Final model
```{r Print Final model}
# Print final model summary
print(summary(reduced_glmer_model))
# Generate confidence intervals for final model
confint.merMod(reduced_glmer_model)

```


